---
title: "add best lineage barcodes"
author: "Andy Sposato"
output:
  pdf_document: default
  html_document: default

---

The purpose of this notebook is to add the best lineage barcode to each cell for which there is a high confidence lineage barcode. We generated a table with cell id and barcode information in the python notebook called find-best-lineage-barcode.ipynb. If you have not yet worked through that script, go back and do so.

```{r load packages, message=FALSE, warning=FALSE}
# load required packages
library(dplyr)
library(Seurat)
library(ggplot2)
library(stringr)
library(cowplot)
```

##### Connect lineage and transcriptome for unedited 

\

```{r load unedited seurat object}
# load the Seurat objects
unedited <- readRDS("data/objects/unedited_w_diffscores.rds")
```

```{r load lineage barcode data for unedited}
# load lineage barcode data
# These tables were written as outputs from find-best-lineage-barcode.ipynb 
un1_data <- read.table('data/tables/unedited1_bestlineagebarcode.tsv', 
                       sep = '\t', header = T)
un2_data <- read.table('data/tables/unedited2_bestlineagebarcode.tsv', 
                       sep = '\t', header = T)
```

```{r unedited metadata}
un_meta <- unedited@meta.data
# add a column that holds cell.id (10X barcode)
un_meta$cell.id <- row.names(un_meta)

# we have a data frame of barcodes associated with unedited1 or unedited2 
# (un1_data and un2_data)
# let's treat these metadata frames separately and then combine them later with rbind
# before adding back as meta data to the Seurat object
un_meta_1 <- subset(un_meta, orig.ident == "mo22_A")
un_meta_2 <- subset(un_meta, orig.ident == "mo22_B")

# get list of cell ids as they appear in each meta data frame
un_cells_1 <- as.data.frame(un_meta_1$cell.id)
names(un_cells_1)[names(un_cells_1) == "un_meta_1$cell.id"] <- "cell.id"
un_cells_2 <- as.data.frame(un_meta_2$cell.id)
names(un_cells_2)[names(un_cells_2) == "un_meta_2$cell.id"] <- "cell.id"
```

```{r add_chars function defined}
source("bin/functions_by_andy.R")
```

```{r add suffix to cell-ids in lineage dataframe for unedited}
# check the cell.id in your seurat objects to verify the correct suffix 
# is being added to your barcode dataframes
head(un1_data$cell.id, n = 2)
un1_data$cell.id <- add_chars(un1_data$cell.id, '-1_1')
head(un1_data$cell.id, n = 2)

head(un2_data$cell.id, n = 2)
un2_data$cell.id <- add_chars(un2_data$cell.id, '-1_2')
head(un2_data$cell.id, n = 2)
```

```{r join transcriptome cell ids with lineage cell ids for unedited}
# using left_join from dplyr maintains the order of the cell_ids according to the Seurat
# object's metadata (stored those in un_cells_1 and un_cells_2) and only uses barcodes 
# for cells which are in the transcriptome (Seurat object)
un_cells_1_barcodes <- left_join(un_cells_1, un1_data)
un_cells_2_barcodes <- left_join(un_cells_2, un2_data)
```

```{r condition subsample identity for unedited}
# add category to differentiate sub samples
# this is likely already stored in your metadata under orig.ident
un_cells_1_barcodes$condition <- "unedited1"
un_cells_2_barcodes$condition <- "unedited2"
```

```{r get new copy of unedited transcriptome cells}
# recopy unedited@meta.data 
# don't want extra cell.id column this time
# subset for just unedited1 (mo22_A) cells 
# and subset for just unedited2 (mo22_B) cells 
Cells = unedited@meta.data
Cells1 <- subset(Cells, orig.ident == "mo22_A")
Cells2 <- subset(Cells, orig.ident == "mo22_B")
```

```{r define the data to be added to transcriptome for unedited1 }
# we define all the lineage barcode information we want to add
barcode_count <- un_cells_1_barcodes[['barcode.count']]
barcode <- un_cells_1_barcodes[['top.barcode']]
site.1 <- un_cells_1_barcodes[['site1']]
site.2 <- un_cells_1_barcodes[['site2']]
site.3 <- un_cells_1_barcodes[['site3']]
site.4 <- un_cells_1_barcodes[['site4']]
site.5 <- un_cells_1_barcodes[['site5']]
site.6 <- un_cells_1_barcodes[['site6']]
site.7 <- un_cells_1_barcodes[['site7']]
site.8 <- un_cells_1_barcodes[['site8']]
site.9 <- un_cells_1_barcodes[['site9']]
site.10 <- un_cells_1_barcodes[['site10']]
```

```{r add barcode data to Cells1 unedited}
# modify Cells1 by adding in those columns
Cells1['barcode.count'] <- barcode_count
Cells1['barcode'] <- barcode
Cells1['site1'] <- site.1
Cells1['site2'] <- site.2
Cells1['site3'] <- site.3
Cells1['site4'] <- site.4
Cells1['site5'] <- site.5
Cells1['site6'] <- site.6
Cells1['site7'] <- site.7
Cells1['site8'] <- site.8
Cells1['site9'] <- site.9
Cells1['site10'] <- site.10
```

```{r subset columns for Seurat metadata integration unedited1}
# now subset the columns that are going into the Seurat object for unedited1
Cells1.info <- subset(Cells1, select = c('barcode.count', 'barcode', 'site1',
                                         'site2','site3', 'site4','site5','site6',
                                         'site7', 'site8','site9', 'site10'))
```

```{r define the data to be added to transcriptome for unedited2}
# we define all the lineage barcode information we want to add for unedited2
barcode_count <- un_cells_2_barcodes[['barcode.count']]
barcode <- un_cells_2_barcodes[['top.barcode']]
site.1 <- un_cells_2_barcodes[['site1']]
site.2 <- un_cells_2_barcodes[['site2']]
site.3 <- un_cells_2_barcodes[['site3']]
site.4 <- un_cells_2_barcodes[['site4']]
site.5 <- un_cells_2_barcodes[['site5']]
site.6 <- un_cells_2_barcodes[['site6']]
site.7 <- un_cells_2_barcodes[['site7']]
site.8 <- un_cells_2_barcodes[['site8']]
site.9 <- un_cells_2_barcodes[['site9']]
site.10 <- un_cells_2_barcodes[['site10']]
```

```{r add barcode data to Cells2 unedited}
# modify Cells2 by adding in those columns
Cells2['barcode.count'] <- barcode_count
Cells2['barcode'] <- barcode
Cells2['site1'] <- site.1
Cells2['site2'] <- site.2
Cells2['site3'] <- site.3
Cells2['site4'] <- site.4
Cells2['site5'] <- site.5
Cells2['site6'] <- site.6
Cells2['site7'] <- site.7
Cells2['site8'] <- site.8
Cells2['site9'] <- site.9
Cells2['site10'] <- site.10
```

```{r subset columns for Seurat metadata integration unedited2}
# now subset the columns that are going into the Seurat object for unedited2
Cells2.info <- subset(Cells2, select = c('barcode.count', 'barcode', 'site1',
                                         'site2','site3', 'site4','site5','site6',
                                         'site7', 'site8','site9', 'site10'))
```

```{r combine unedited dataframes}
# stack these two data frames on top of each other with rbind()
Cells.both_un <- rbind(Cells1.info, Cells2.info)

# for plotting purposes later, we want to change 'NA' barcode.count values to 0
Cells.both_un$barcode.count[is.na(Cells.both_un$barcode.count)] <- 0
```

```{r add lineage barcode data to unedited transcriptome}
# add back to Seurat object unedited
unedited <- AddMetaData(unedited, Cells.both_un)
```

```{r save unedited object with lineage information}
saveRDS(unedited, "data/objects/unedited_w_diffscores_lineage.rds")
```

##### Connect lineage and transcriptome for edited

\

```{r load seurat object for edited}
# load the Seurat objects
edited <- readRDS("data/objects/edited_w_diffscores.rds")
```

```{r load lineage barcode data for edited}
# load lineage barcode data
ed1_data <- read.table('data/tables/inj_heat1_bestlineagebarcode.tsv', 
                       sep = '\t', header = T)
ed2_data <- read.table('data/tables/inj_heat2_bestlineagebarcode.tsv', 
                       sep = '\t', header = T)
```

```{r edited metadata}
ed_meta <- edited@meta.data
# add a column that holds cell.id (10X barcode)
ed_meta$cell.id <- row.names(ed_meta)

ed_meta_1 <- subset(ed_meta, orig.ident == "mo20_A")
ed_meta_2 <- subset(ed_meta, orig.ident == "mo20_B")

# get list of cell ids as they appear in each meta data frame
ed_cells_1 <- as.data.frame(ed_meta_1$cell.id)
names(ed_cells_1)[names(ed_cells_1) == "ed_meta_1$cell.id"] <- "cell.id"
ed_cells_2 <- as.data.frame(ed_meta_2$cell.id)
names(ed_cells_2)[names(ed_cells_2) == "ed_meta_2$cell.id"] <- "cell.id"
```

```{r add suffix to edited cell-ids in lineage dataframe}
# we added the add_chars function earlier
head(ed1_data$cell.id, n = 2)
ed1_data$cell.id <- add_chars(ed1_data$cell.id, '-1')
head(ed1_data$cell.id, n = 2)

head(ed2_data$cell.id, n = 2)
ed2_data$cell.id <- add_chars(ed2_data$cell.id, '-1')
head(ed2_data$cell.id, n = 2)
```

```{r join edited transcriptome cell ids with lineage cell ids}
ed_cells_1_barcodes <- left_join(ed_cells_1, ed1_data)
ed_cells_2_barcodes <- left_join(ed_cells_2, ed2_data)
```

```{r condition subsample identity for edited}
# this is really just differentiating technical replicates
# same biological samples, different channels of 10X
ed_cells_1_barcodes$condition <- "edited1"
ed_cells_2_barcodes$condition <- "edited2"
```

```{r get new copy of edited transcriptome cells}
# recopy edited@meta.data 
# don't want extra cell.id column this time
# subset for just edited1 cells 
# also subset for just edited2 cells 
Cells = edited@meta.data
Cells1 <- subset(Cells, orig.ident == "mo20_A")
Cells2 <- subset(Cells, orig.ident == "mo20_B")
```

```{r define the data to be added to transcriptome for edited1}
# we define all the lineage barcode information we want to add
barcode_count <- ed_cells_1_barcodes[['barcode.count']]
barcode <- ed_cells_1_barcodes[['top.barcode']]
site.1 <- ed_cells_1_barcodes[['site1']]
site.2 <- ed_cells_1_barcodes[['site2']]
site.3 <- ed_cells_1_barcodes[['site3']]
site.4 <- ed_cells_1_barcodes[['site4']]
site.5 <- ed_cells_1_barcodes[['site5']]
site.6 <- ed_cells_1_barcodes[['site6']]
site.7 <- ed_cells_1_barcodes[['site7']]
site.8 <- ed_cells_1_barcodes[['site8']]
site.9 <- ed_cells_1_barcodes[['site9']]
site.10 <- ed_cells_1_barcodes[['site10']]
```

```{r add barcode data to Cells1 edited}
# modify Cells1 by adding in those columns
Cells1['barcode.count'] <- barcode_count
Cells1['barcode'] <- barcode
Cells1['site1'] <- site.1
Cells1['site2'] <- site.2
Cells1['site3'] <- site.3
Cells1['site4'] <- site.4
Cells1['site5'] <- site.5
Cells1['site6'] <- site.6
Cells1['site7'] <- site.7
Cells1['site8'] <- site.8
Cells1['site9'] <- site.9
Cells1['site10'] <- site.10
```

```{r subset columns for Seurat metadata integration for edited1}
# now subset the columns that are going into the Seurat object for edited1
Cells1.info <- subset(Cells1, select = c('barcode.count', 'barcode', 'site1',
                                         'site2','site3', 'site4','site5','site6',
                                         'site7', 'site8','site9', 'site10'))
```

```{r define the data to be added to transcriptome for edited2}
# we define all the lineage barcode information we want to add for edited2
barcode_count <- ed_cells_2_barcodes[['barcode.count']]
barcode <- ed_cells_2_barcodes[['top.barcode']]
site.1 <- ed_cells_2_barcodes[['site1']]
site.2 <- ed_cells_2_barcodes[['site2']]
site.3 <- ed_cells_2_barcodes[['site3']]
site.4 <- ed_cells_2_barcodes[['site4']]
site.5 <- ed_cells_2_barcodes[['site5']]
site.6 <- ed_cells_2_barcodes[['site6']]
site.7 <- ed_cells_2_barcodes[['site7']]
site.8 <- ed_cells_2_barcodes[['site8']]
site.9 <- ed_cells_2_barcodes[['site9']]
site.10 <- ed_cells_2_barcodes[['site10']]
```

```{r add barcode data to Cells2 edited}
# modify Cells2 by adding in those columns
Cells2['barcode.count'] <- barcode_count
Cells2['barcode'] <- barcode
Cells2['site1'] <- site.1
Cells2['site2'] <- site.2
Cells2['site3'] <- site.3
Cells2['site4'] <- site.4
Cells2['site5'] <- site.5
Cells2['site6'] <- site.6
Cells2['site7'] <- site.7
Cells2['site8'] <- site.8
Cells2['site9'] <- site.9
Cells2['site10'] <- site.10
```

```{r subset columns for Seurat metadata integration for edited2}
# now subset the columns that are going into the Seurat object for edited2
Cells2.info <- subset(Cells2, select = c('barcode.count', 'barcode', 'site1',
                                         'site2','site3', 'site4','site5','site6',
                                         'site7', 'site8','site9', 'site10'))
```

```{r combine edited dataframes}
# stack these two data frames on top of each other with rbind()
Cells.both_ed <- rbind(Cells1.info, Cells2.info)

# for plotting purposes later, we want to change 'NA' barcode.count labels to 0
Cells.both_ed$barcode.count[is.na(Cells.both_ed$barcode.count)] <- 0
```

```{r add lineage barcode data to edited transcriptome}
# add back to Seurat object unedited
edited <- AddMetaData(edited, Cells.both_ed)
```

```{r save edited object}
saveRDS(edited, "data/objects/edited_w_diffscores_lineage.rds")
```
