---
title: "scoring cells by differentiation"
author: "Andy Sposato"
output: html_document
---

#### The purpose of this notebook is to add differentation scores to Seurat objects. We do this by converting Seurat objects to URD objects, running URD, and adding the pseudotime information to a Seurat object's metadata as diff.score. 

##### Generate the Seurat objects
In this example, I'll be using an unedited sample (yr1.6 or 20 mo.) and a double-edited sample (yr1.8 or 22 mo.)

If you have already generated objects, skip to Run URD section. 
You have likely already done this as you'll want to see quality of your atlas before you proceed with amplifying lineage barcodes from 10X cDNA.

```{r load packages, message=FALSE, warning=FALSE}
# load required packages
library(dplyr)
library(Seurat)
# source("https://raw.githubusercontent.com/farrellja/URD/master/URD-Install.R") # uncomment if you need to install URD
suppressPackageStartupMessages(library(rgl))
## Warning: package 'rgl' was built under R version 3.4.3
suppressPackageStartupMessages(library(URD))
## Warning: package 'Matrix' was built under R version 3.4.2
knitr::opts_chunk$set(echo = TRUE)
rgl::setupKnitr()
library(scales)
```

```{r make 20 mo. object, message=FALSE, warning=FALSE}
yr1.6_A.data <- Read10X(data.dir = 'data/18827X2/filtered_feature_bc_matrix/')
yr1.6_A <- CreateSeuratObject(counts = yr1.6_A.data, project = 'yr1.6_A', min.cells = 3)
yr1.6_A[['percent.mt']] <- PercentageFeatureSet(yr1.6_A, pattern = "mt-")
yr1.6_A$sub.sample <- "20mo"
yr1.6_A$age <- "20 mo."
yr1.6_A$condition <- "unedited"
yr1.6_A <- subset(yr1.6_A, subset = nFeature_RNA > 200 & nFeature_RNA < 4999)
yr1.6_A <- subset(yr1.6_A, subset = percent.mt < 5)
yr1.6_A<- NormalizeData(yr1.6_A)
yr1.6_A <- FindVariableFeatures(yr1.6_A)
yr1.6_A <- ScaleData(yr1.6_A)
yr1.6_A <- RunPCA(yr1.6_A)

yr1.6_B.data <- Read10X(data.dir = 'data/18827X4/filtered_feature_bc_matrix/')
yr1.6_B <- CreateSeuratObject(counts = yr1.6_B.data, project = 'yr1.6_B', min.cells = 3)
yr1.6_B <- subset(yr1.6_B, subset = nFeature_RNA > 200 & nFeature_RNA < 6373)
yr1.6_B[['percent.mt']] <- PercentageFeatureSet(yr1.6_B, pattern = "mt-")
yr1.6_B$sub.sample <- "20mo"
yr1.6_B$age <- "20 mo."
yr1.6_B$condition <- "unedited"
yr1.6_B <- subset(yr1.6_B, subset = percent.mt < 5)
yr1.6_B <- NormalizeData(yr1.6_B)
yr1.6_B <- FindVariableFeatures(yr1.6_B)
yr1.6_B <- ScaleData(yr1.6_B)
yr1.6_B <- RunPCA(yr1.6_B)

yr1.6.anchors <- FindIntegrationAnchors(object.list = list(yr1.6_A, yr1.6_B), reduction = "cca", dims = 1:10)
unedited <- IntegrateData(anchorset = yr1.6.anchors, dims = 1:10)
DefaultAssay(unedited) <- "integrated"
unedited <- ScaleData(unedited, assay = "integrated")
unedited <- ScaleData(unedited, assay = "RNA")
unedited <- RunPCA(unedited, features = VariableFeatures(object = unedited))
unedited <- RunUMAP(unedited, reduction = 'pca', dims = 1:10)
unedited <- FindNeighbors(unedited, reduction = 'pca', dims = 1:10)
unedited <- FindClusters(unedited, resolution = 0.5)

rm(yr1.6_A.data, yr1.6_B.data, yr1.6_A, yr1.6_B, yr1.6.anchors)
```

```{r repeat for 22 mo., message=FALSE, warning=FALSE}
yr1.8_A.data <- Read10X(data.dir = 'data/18827X1/filtered_feature_bc_matrix/')
# create Seurat object
yr1.8_A <- CreateSeuratObject(counts = yr1.8_A.data, project = 'yr1.8_A', min.cells = 3)
yr1.8_A[['percent.mt']] <- PercentageFeatureSet(yr1.8_A, pattern = "mt-")
yr1.8_A$sub.sample <- "22mo" # for comparison of data set X3 vs X4
yr1.8_A$age <- "22 mo."
yr1.8_A$condition <- "edited"
yr1.8_A <- subset(yr1.8_A, subset = nFeature_RNA > 200 & nFeature_RNA < 6962)
yr1.8_A <- subset(yr1.8_A, subset = percent.mt < 5)
yr1.8_A <- NormalizeData(yr1.8_A)
yr1.8_A <- FindVariableFeatures(yr1.8_A)
yr1.8_A <- ScaleData(yr1.8_A)
yr1.8_A <- RunPCA(yr1.8_A)

yr1.8_B.data <- Read10X(data.dir = 'data/18827X3/filtered_feature_bc_matrix/')
yr1.8_B <- CreateSeuratObject(counts = yr1.8_B.data, project = 'yr1.8_B', min.cells = 3)
yr1.8_B[['percent.mt']] <- PercentageFeatureSet(yr1.8_B, pattern = "mt-")
yr1.8_B$sub.sample <- "22mo"
yr1.8_B$age <- "22 mo."
yr1.8_B$condition <- "edited"
yr1.8_B <- subset(yr1.8_B, subset = nFeature_RNA > 200 & nFeature_RNA < 7058)
yr1.8_B <- subset(yr1.8_B, subset = percent.mt < 5)
yr1.8_B <- NormalizeData(yr1.8_B)
yr1.8_B <- FindVariableFeatures(yr1.8_B)
yr1.8_B <- ScaleData(yr1.8_B)
yr1.8_B <- RunPCA(yr1.8_B)

yr1.8.anchors <- FindIntegrationAnchors(object.list = list(yr1.8_A, yr1.8_B), reduction = "cca", dims = 1:15)
edited <- IntegrateData(anchorset = yr1.8.anchors, dims = 1:15)
DefaultAssay(edited) <- "integrated"
edited <- ScaleData(edited, assay = "integrated")
edited <- ScaleData(edited, assay = "RNA")
edited <- RunPCA(edited, features = VariableFeatures(object = edited))
edited <- RunUMAP(edited, reduction = 'pca', dims = 1:15)
edited <- FindNeighbors(edited, reduction = 'pca', dims = 1:15)
edited <- FindClusters(edited, resolution = 0.5)

rm(yr1.8_A.data, yr1.8_B.data, yr1.8_A, yr1.8_B, yr1.8.anchors)
```

##### Run URD 
URD is a pseudotime approach that assigns a differentiation score to each individual cell in your object. The more cells you have, the longer URD takes to run. You may consider running this program as a scheduled job through CHPC. 

```{r load objects}
unedited <- readRDS("data/objects/mo22.rds")
edited <- readRDS("data/objects/mo20.rds")
```

We're looking at ddx4 as a marker for spermatogonia. The most undifferentiated cells will express this marker well and help us determine where the "beginning" of differentiation is. This will be important for setting the root cells when running URD.
```{r plot ddx4 to see directionality}
DimPlot(unedited, label = T, label.size = 5)
FeaturePlot(unedited, "dazl")
DimPlot(edited, label = T, label.size = 5)
FeaturePlot(edited, "dazl")
```

This seuratV3ToURD function was written by Jeff Farrell. It converts a Seurat object to an URD object. Feel free to open it up in a text editor to review it. 
```{r convert Seurat object to URD object}
source("bin/SeuratV3ToURD.R")
unedited_URD <- seuratV3ToURD(unedited)
edited_URD <- seuratV3ToURD(edited)
```

```{r run URD on unedited}
clusters <- as.factor(unedited_URD@group.ids$seurat_clusters)
unedited_URD@meta$seurat_clusters <- clusters

var.genes <- findVariableGenes(unedited_URD, set.object.var.genes=F, diffCV.cutoff=0.3, mean.min=.005, mean.max=100, main.use="", do.plot = F)
unedited_URD@var.genes <- var.genes

unedited_URD <- calcDM(unedited_URD, knn = 57)

# the root cells for the forward calculation should be your spermatogonia clusters
root.cells <- cellsInCluster(unedited_URD, "seurat_clusters", c("4", "8"))
unedited_URD.floods <- floodPseudotime(unedited_URD, root.cells = root.cells, n=20, minimum.cells.flooded = 2, verbose=F)
unedited_URD <- floodPseudotimeProcess(unedited_URD, unedited_URD.floods, floods.name="pseudotime.fwd")

# the root cells for the reverse calculation should be the most mature spermatids cluster
root.cells <- cellsInCluster(unedited_URD, "seurat_clusters", c("7"))
unedited_URD.floods <- floodPseudotime(unedited_URD, root.cells = root.cells, n=20, minimum.cells.flooded = 2, verbose=F)
unedited_URD <- floodPseudotimeProcess(unedited_URD, unedited_URD.floods, floods.name="pseudotime.rev")

# normalize values 
unedited_URD@pseudotime$pseudotime.fwd.norm <- unedited_URD@pseudotime$pseudotime.fwd/max(unedited_URD@pseudotime$pseudotime.fwd)
unedited_URD@pseudotime$pseudotime.rev.norm <- 1 - unedited_URD@pseudotime$pseudotime.rev/max(unedited_URD@pseudotime$pseudotime.rev)

# take the average of the normalized forward and reverse
unedited_URD@pseudotime$pseudotime <- rowMeans(unedited_URD@pseudotime[,c("pseudotime.rev.norm", "pseudotime.fwd.norm")])

# rescale so differentiation scores are on a 0 to 1 range
unedited_URD@pseudotime$pseudotime.scaled <- rescale(unedited_URD@pseudotime$pseudotime, to = c(0,1))

# save the object - later we'll pull pseudotime.scaled from it for transfer to the Seurat object
saveRDS(unedited_URD, "data/objects/unedited_URD.rds")
rm(unedited_URD.floods)
```

```{r run URD on edited}
clusters <- as.factor(edited_URD@group.ids$seurat_clusters)
edited_URD@meta$seurat_clusters <- clusters

var.genes <- findVariableGenes(edited_URD, set.object.var.genes=F, diffCV.cutoff=0.3, mean.min=.005, mean.max=100, main.use="", do.plot = F)
edited_URD@var.genes <- var.genes

edited_URD <- calcDM(edited_URD, knn = 51)

# the root cells for the forward calculation should be your spermatogonia clusters
root.cells <- cellsInCluster(edited_URD, "seurat_clusters", c("9", "10"))
edited_URD.floods <- floodPseudotime(edited_URD, root.cells = root.cells, n=20, minimum.cells.flooded = 2, verbose=F)
edited_URD <- floodPseudotimeProcess(edited_URD, edited_URD.floods, floods.name="pseudotime.fwd")

# the root cells for the reverse calculation should be the most mature spermatids cluster
root.cells <- cellsInCluster(edited_URD, "seurat_clusters", c("7"))
edited_URD.floods <- floodPseudotime(edited_URD, root.cells = root.cells, n=20, minimum.cells.flooded = 2, verbose=F)
edited_URD <- floodPseudotimeProcess(edited_URD, edited_URD.floods, floods.name="pseudotime.rev")

# normalize values 
edited_URD@pseudotime$pseudotime.fwd.norm <- edited_URD@pseudotime$pseudotime.fwd/max(edited_URD@pseudotime$pseudotime.fwd)
edited_URD@pseudotime$pseudotime.rev.norm <- 1 -  edited_URD@pseudotime$pseudotime.rev/max(edited_URD@pseudotime$pseudotime.rev)

# take the average of the normalized forward and reverse
edited_URD@pseudotime$pseudotime <- rowMeans(edited_URD@pseudotime[,c("pseudotime.rev.norm", "pseudotime.fwd.norm")])

# rescale so differentiation scores are on a 0 to 1 range
edited_URD@pseudotime$pseudotime.scaled <- rescale(edited_URD@pseudotime$pseudotime, to = c(0,1))

# save the object - later we'll pull pseudotime.scaled from it for transfer to the Seurat object
saveRDS(edited_URD, "data/objects/edited_URD.rds")
rm(edited_URD.floods)
```

```{r check pseudotime for unedited}
DimPlot(unedited, group.by = "seurat_clusters", label = T)
FeaturePlot(unedited, "rna_ddx4")
FeaturePlot(unedited, "rna_gsdf")
FeaturePlot(unedited, "rna_tssk6")
plotDim(unedited_URD, "pseudotime.fwd")
plotDim(unedited_URD, "pseudotime.rev")
plotDim(unedited_URD, "pseudotime")
plotDim(unedited_URD, "pseudotime.scaled")
```

```{r check pseudotime for edited}
DimPlot(edited, group.by = "seurat_clusters", label = T)
FeaturePlot(edited, "rna_ddx4")
FeaturePlot(edited, "rna_gsdf")
FeaturePlot(edited, "rna_tssk6")
plotDim(edited_URD, "pseudotime.fwd")
plotDim(edited_URD, "pseudotime.rev")
plotDim(edited_URD, "pseudotime")
plotDim(edited_URD, "pseudotime.scaled")
```

##### Add pseudotime information from URD objects to Seurat objects
```{r add differentiation scores to objects}
unedited_diffs <- as.vector(unedited_URD@pseudotime$pseudotime.scaled)
edited_diffs <- as.vector(edited_URD@pseudotime$pseudotime.scaled)

unedited@meta.data$diff.score <- unedited_diffs
edited@meta.data$diff.score <- edited_diffs

saveRDS(unedited, "data/objects/unedited_w_diffscores.rds")
saveRDS(edited, "data/objects/edited_w_diffscores.rds")
```

```{r check diff.scores were added}
FeaturePlot(unedited, "diff.score")
FeaturePlot(edited, "diff.score")
```
