---
title: "visualize lineage barcodes"
author: "Andy Sposato"
output: html_document
---

##### The purpose of this notebook is to visualize lineage barcodes individually on UMAP and in a line graph. 
For individual barcodes on UMAP, we can generate UMAP images where all cells with the same lineage barcode are highlighted. 
For the line graph, we can plot normalized barcode abundance across differentiation within the object. 

```{r load packages, message=FALSE, warning=FALSE}
# load required packages
library(dplyr)
library(Seurat)
library(ggplot2)
library(stringr)
library(cowplot)
library(reshape2)
source("bin/plot_barcode_on_UMAP.R")
source("bin/functions_by_andy.R") 
```

##### Plotting individual barcodes on UMAP

```{r load your objects}
edited <- readRDS("data/objects/edited_w_diffscores_lineage.rds")
unedited <- readRDS("data/objects/unedited_w_diffscores_lineage.rds")
```

```{r examples}
plot_barcode("84D+45_84D+45_84D+45_84D+45_NONE_NONE_NONE_NONE_NONE_NONE", edited, "firebrick3")
plot_barcode("NONE_NONE_NONE_NONE_NONE_NONE_NONE_NONE_NONE_NONE", unedited, "dodgerblue")
```
##### Generating a line graph for all the barcodes in your object

```{r load your objects}
edited <- readRDS("data/objects/edited_w_diffscores_lineage.rds")
unedited <- readRDS("data/objects/unedited_w_diffscores_lineage.rds")
```

###### edited data 
```{r write diffscores to a file to convert to diffranges in python}
meta <- edited@meta.data
meta$cell.id <- row.names(meta)
meta <- as.data.frame(cbind(meta$cell.id, meta$diff.score, meta$barcode, meta$site1, meta$site2, meta$site3, meta$site4, meta$site5, meta$site6, meta$site7, meta$site8, meta$site9, meta$site10))
write.table(meta, "data/tables/edited_diffscores.tsv", sep = "\t", col.names = F, row.names = F, quote = F)
```

```{python save data folder string}
data_folder = "./data/tables/"
```

```{python create cells_ranges_dict}
cells_ranges_dict= {}
file = open(data_folder + 'edited_diffscores.tsv', "r")
for line in file.readlines(): 
    line = line.split("\t")
    cellid = line[0].rstrip("\t")
    diff = float(line[1].rstrip("\t"))
    barcode = line[2].rstrip("\t")
    site1 = line[3].rstrip("\t")
    site2 = line[4].rstrip("\t")
    site3 = line[5].rstrip("\t")
    site4 = line[6].rstrip("\t")
    site5 = line[7].rstrip("\t")
    site6 = line[8].rstrip("\t")
    site7 = line[9].rstrip("\t")
    site8 = line[10].rstrip("\t")
    site9 = line[11].rstrip("\t")
    site10 = line[12].rstrip("\n\t")
    if diff <= 0.1: 
        cells_ranges_dict[cellid] = ['0-0.1', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.1 and diff <= 0.2: 
        cells_ranges_dict[cellid] = ['0.1-0.2', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.2 and diff <= 0.3: 
        cells_ranges_dict[cellid] = ['0.2-0.3', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.3 and diff <= 0.4: 
        cells_ranges_dict[cellid] = ['0.3-0.4', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.4 and diff <= 0.5: 
        cells_ranges_dict[cellid] = ['0.4-0.5', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.5 and diff <= 0.6: 
        cells_ranges_dict[cellid] = ['0.5-0.6', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.6 and diff <= 0.7: 
        cells_ranges_dict[cellid] = ['0.6-0.7', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.7 and diff <= 0.8: 
        cells_ranges_dict[cellid] = ['0.7-0.8', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.8 and diff <= 0.9: 
        cells_ranges_dict[cellid] = ['0.8-0.9', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.9: 
        cells_ranges_dict[cellid] = ['0.9-1.0', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
file.close()
```

```{python write diffranges to a table, include=FALSE}
output = open(data_folder+"edited_diffranges.tsv", 'w')
output.write("cell.id\t" + "diff.range\t" + "barcode\t" + "site1\t" + "site2\t" + "site3\t" + "site4\t" + "site5\t" + "site6\t" + "site7\t" + "site8\t" + "site9\t" + "site10\n")
for cell, barcode_info in cells_ranges_dict.items(): 
    output.write(cell + '\t')
    for info in barcode_info[0:-1]: 
        output.write(info + '\t')
    output.write(barcode_info[-1])
    output.write('\n')
output.close()
```

```{r add diff.range column to transcriptome metadata}
edited_diffranges <- read.table('data/tables/edited_diffranges.tsv', sep = '\t', header = T)
edited <- AddMetaData(edited, as.list(subset(edited_diffranges, select = c('diff.range'))), col.name = 'diff.range')
```

```{r}
unique_edited_barcodes <- as.data.frame(unique(edited@meta.data$barcode))
write.table(unique_edited_barcodes, "data/tables/unique_edited_barcodes.tsv", sep = "\t", col.names = F, row.names = F, quote = F)

```

```{python collect unique barcodes from edited testes}
unique_edited_barcodes = open(data_folder + "unique_edited_barcodes.tsv", "r")
barcodes = []
for barcode in unique_edited_barcodes: 
  barcode_string = barcode.strip("\n")
  if barcode_string != "NA":
    barcodes.append(barcode_string)
unique_edited_barcodes.close()
```

```{python }
barcodes_dict = {}
for barcode in barcodes: 
    barcodes_dict[str(barcode)] = {'0-0.1': [], '0.1-0.2': [], '0.2-0.3': [], '0.3-0.4': [], '0.4-0.5': [], '0.5-0.6': [], '0.6-0.7': [], '0.7-0.8': [], '0.8-0.9': [], '0.9-1': []}
```

```{python}
barcodes_dict['1I+45+C_15D+66_NONE_9D+123_NONE_NONE_NONE_NONE_NONE_NONE']
```

```{python}
file = open(data_folder + "edited_diffranges.tsv", "r")
file.readline()
for line in file.readlines(): 
    line = line.split("\t")
    cell = line[0].strip("\t")
    diff_range = line[1].strip("\t")
    barcode = line[2].rstrip("\t")
    if barcode in barcodes_dict.keys(): 
        if diff_range in barcodes_dict[barcode].keys(): 
          barcodes_dict[barcode][diff_range].append(cell)
file.close()
```

```{python}
barcodes_dict['1I+45+C_15D+66_NONE_9D+123_NONE_NONE_NONE_NONE_NONE_NONE']
```

```{python}
for ranges_dictionary in barcodes_dict.values(): 
  for key, value in ranges_dictionary.items(): 
    cell_count = len(value)
    ranges_dictionary[key].append(str(cell_count))
```

```{python}
barcodes_dict['1I+45+C_15D+66_NONE_9D+123_NONE_NONE_NONE_NONE_NONE_NONE']
```

```{python}
output = open(data_folder + 'edited_diffranges_cellcounts.tsv', 'w')
output.write('barcode\t'+'0-0.1'+'\t'+'0.1-0.2'+'\t'+'0.2-0.3'+'\t'+'0.3-0.4'+'\t'+'0.4-0.5'+'\t'+
             '0.5-0.6'+'\t'+'0.6-0.7'+'\t'+'0.7-0.8'+'\t'+'0.8-0.9'+'\t'+'0.9-1'+'\n')
for barcode in barcodes_dict.keys():  
        output.write(barcode+
                 '\t'+barcodes_dict[barcode]['0-0.1'][-1]+
                 '\t'+barcodes_dict[barcode]['0.1-0.2'][-1]+
                 '\t'+barcodes_dict[barcode]['0.2-0.3'][-1]+
                 '\t'+barcodes_dict[barcode]['0.3-0.4'][-1]+
                 '\t'+barcodes_dict[barcode]['0.4-0.5'][-1]+
                 '\t'+barcodes_dict[barcode]['0.5-0.6'][-1]+
                 '\t'+barcodes_dict[barcode]['0.6-0.7'][-1]+
                 '\t'+barcodes_dict[barcode]['0.7-0.8'][-1]+
                 '\t'+barcodes_dict[barcode]['0.8-0.9'][-1]+
                 '\t'+barcodes_dict[barcode]['0.9-1'][-1]+'\n')
output.close()
```

```{r make clonality dataframe}
clonality <- read.table("data/tables/edited_diffranges_cellcounts.tsv", sep = "\t", header = T)
clonality$X0.0.1 <- as.vector(sapply(clonality$X0.0.1,normalize,length(subset(edited@meta.data, barcode.count != 0 & diff.range == '0-0.1')$barcode)))
clonality$X0.1.0.2 <- as.vector(sapply(clonality$X0.1.0.2,normalize,length(subset(edited@meta.data, barcode.count != 0 & diff.range == '0.1-0.2')$barcode)))
clonality$X0.2.0.3 <- as.vector(sapply(clonality$X0.2.0.3,normalize,length(subset(edited@meta.data, barcode.count != 0 & diff.range == '0.2-0.3')$barcode)))
clonality$X0.3.0.4 <- as.vector(sapply(clonality$X0.3.0.4,normalize,length(subset(edited@meta.data, barcode.count != 0 & diff.range == '0.3-0.4')$barcode)))
clonality$X0.4.0.5 <- as.vector(sapply(clonality$X0.4.0.5,normalize,length(subset(edited@meta.data, barcode.count != 0 & diff.range == '0.4-0.5')$barcode)))
clonality$X0.5.0.6 <- as.vector(sapply(clonality$X0.5.0.6,normalize,length(subset(edited@meta.data, barcode.count != 0 & diff.range == '0.5-0.6')$barcode)))
clonality$X0.6.0.7 <- as.vector(sapply(clonality$X0.6.0.7,normalize,length(subset(edited@meta.data, barcode.count != 0 & diff.range == '0.6-0.7')$barcode)))
clonality$X0.7.0.8 <- as.vector(sapply(clonality$X0.7.0.8,normalize,length(subset(edited@meta.data, barcode.count != 0 & diff.range == '0.7-0.8')$barcode)))
clonality$X0.8.0.9 <- as.vector(sapply(clonality$X0.8.0.9,normalize,length(subset(edited@meta.data, barcode.count != 0 & diff.range == '0.8-0.9')$barcode)))
clonality$X0.9.1<- as.vector(sapply(clonality$X0.9.1,normalize,length(subset(edited@meta.data, barcode.count != 0 & diff.range == '0.9-1.0')$barcode)))
names(clonality)[names(clonality) == "X0.0.1"] <- "0 - 0.1"
names(clonality)[names(clonality) == "X0.1.0.2"] <- "0.1 - 0.2"
names(clonality)[names(clonality) == "X0.2.0.3"] <- "0.2 - 0.3"
names(clonality)[names(clonality) == "X0.3.0.4"] <- "0.3 - 0.4"
names(clonality)[names(clonality) == "X0.4.0.5"] <- "0.4- 0.5"
names(clonality)[names(clonality) == "X0.5.0.6"] <- "0.5 - 0.6"
names(clonality)[names(clonality) == "X0.6.0.7"] <- "0.6 - 0.7"
names(clonality)[names(clonality) == "X0.7.0.8"] <- "0.7 - 0.8"
names(clonality)[names(clonality) == "X0.8.0.9"] <- "0.8 - 0.9"
names(clonality)[names(clonality) == "X0.9.1"] <- "0.9 - 1"
```

```{r plot line graph}
clonality_df <- melt(clonality, id.vars="barcode", value.name="value", variable.name= c("state"))
ggplot(data=clonality_df, aes(x=state, y=value, group = barcode, colour = barcode)) +
  #geom_point() + 
  geom_line(size = 1) +
  theme_bw() +
  theme(text = element_text(family = "Arial")) +
  theme(axis.text.y = element_text(size = 10)) +
  theme(axis.text.x = element_text(size = 10, angle = 25, hjust = 1)) +
  theme(axis.title = element_text(size = 15)) +
  theme(plot.title = element_text(size = 20)) +
  ggtitle("lineage barcode abundance across differentiation") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("differentiation ranges") + 
  ylab("normalized percent") + ylim(0,50) + 
  scale_color_manual(values=rainbow(14)) + theme(legend.position = "none")
```


###### unedited data 
This will be less exciting since all barcodes found should be the unedited barcode. 
But it serves as a good check that it's working right. We should see the unedited barcode at 100% abundance across all differentiation ranges except 0.9-1 because we don't actually have barcoded cells in that range. 
```{r write diffscores to a file to convert to diffranges in python}
meta <- unedited@meta.data
meta$cell.id <- row.names(meta)
meta <- as.data.frame(cbind(meta$cell.id, meta$diff.score, meta$barcode, meta$site1, meta$site2, meta$site3, meta$site4, meta$site5, meta$site6, meta$site7, meta$site8, meta$site9, meta$site10))
write.table(meta, "data/tables/unedited_diffscores.tsv", sep = "\t", col.names = F, row.names = F, quote = F)
```

```{python save data folder string}
data_folder = "./data/tables/"
```

```{python create cells_ranges_dict}
cells_ranges_dict= {}
file = open(data_folder + 'unedited_diffscores.tsv', "r")
for line in file.readlines(): 
    line = line.split("\t")
    cellid = line[0].rstrip("\t")
    diff = float(line[1].rstrip("\t"))
    barcode = line[2].rstrip("\t")
    site1 = line[3].rstrip("\t")
    site2 = line[4].rstrip("\t")
    site3 = line[5].rstrip("\t")
    site4 = line[6].rstrip("\t")
    site5 = line[7].rstrip("\t")
    site6 = line[8].rstrip("\t")
    site7 = line[9].rstrip("\t")
    site8 = line[10].rstrip("\t")
    site9 = line[11].rstrip("\t")
    site10 = line[12].rstrip("\n\t")
    if diff <= 0.1: 
        cells_ranges_dict[cellid] = ['0-0.1', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.1 and diff <= 0.2: 
        cells_ranges_dict[cellid] = ['0.1-0.2', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.2 and diff <= 0.3: 
        cells_ranges_dict[cellid] = ['0.2-0.3', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.3 and diff <= 0.4: 
        cells_ranges_dict[cellid] = ['0.3-0.4', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.4 and diff <= 0.5: 
        cells_ranges_dict[cellid] = ['0.4-0.5', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.5 and diff <= 0.6: 
        cells_ranges_dict[cellid] = ['0.5-0.6', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.6 and diff <= 0.7: 
        cells_ranges_dict[cellid] = ['0.6-0.7', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.7 and diff <= 0.8: 
        cells_ranges_dict[cellid] = ['0.7-0.8', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.8 and diff <= 0.9: 
        cells_ranges_dict[cellid] = ['0.8-0.9', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
    elif diff > 0.9: 
        cells_ranges_dict[cellid] = ['0.9-1.0', barcode, site1, site2, site3, site4, site5, site6, site7, site8, site9, site10]
file.close()
```

```{python write diffranges to a table, include=FALSE}
output = open(data_folder+"unedited_diffranges.tsv", 'w')
output.write("cell.id\t" + "diff.range\t" + "barcode\t" + "site1\t" + "site2\t" + "site3\t" + "site4\t" + "site5\t" + "site6\t" + "site7\t" + "site8\t" + "site9\t" + "site10\n")
for cell, barcode_info in cells_ranges_dict.items(): 
    output.write(cell + '\t')
    for info in barcode_info[0:-1]: 
        output.write(info + '\t')
    output.write(barcode_info[-1])
    output.write('\n')
output.close()
```

```{r add diff.range column to transcriptome metadata}
unedited_diffranges <- read.table('data/tables/unedited_diffranges.tsv', sep = '\t', header = T)
unedited <- AddMetaData(unedited, as.list(subset(unedited_diffranges, select = c('diff.range'))), col.name = 'diff.range')
```

```{r}
unique_unedited_barcodes <- as.data.frame(unique(unedited@meta.data$barcode))
write.table(unique_unedited_barcodes, "data/tables/unique_unedited_barcodes.tsv", sep = "\t", col.names = F, row.names = F, quote = F)

```

```{python collect unique barcodes from edited testes}
unique_unedited_barcodes = open(data_folder + "unique_unedited_barcodes.tsv", "r")
barcodes = []
for barcode in unique_unedited_barcodes: 
  barcode_string = barcode.strip("\n")
  if barcode_string != "NA":
    barcodes.append(barcode_string)
unique_unedited_barcodes.close()
```

```{python }
barcodes_dict = {}
for barcode in barcodes: 
    barcodes_dict[str(barcode)] = {'0-0.1': [], '0.1-0.2': [], '0.2-0.3': [], '0.3-0.4': [], '0.4-0.5': [], '0.5-0.6': [], '0.6-0.7': [], '0.7-0.8': [], '0.8-0.9': [], '0.9-1': []}
```


```{python}
barcodes_dict['NONE_NONE_NONE_NONE_NONE_NONE_NONE_NONE_NONE_NONE']
```

```{python}
file = open(data_folder + "unedited_diffranges.tsv", "r")
file.readline()
for line in file.readlines(): 
    line = line.split("\t")
    cell = line[0].strip("\t")
    diff_range = line[1].strip("\t")
    barcode = line[2].rstrip("\t")
    if barcode in barcodes_dict.keys(): 
        if diff_range in barcodes_dict[barcode].keys(): 
          barcodes_dict[barcode][diff_range].append(cell)
file.close()
```

```{python}
# checking this worked right
print(barcodes_dict['NONE_NONE_NONE_NONE_NONE_NONE_NONE_NONE_NONE_NONE']['0.8-0.9'])
```

```{python}
for ranges_dictionary in barcodes_dict.values(): 
  for key, value in ranges_dictionary.items(): 
    cell_count = len(value)
    ranges_dictionary[key].append(str(cell_count))
```

```{python}
output = open(data_folder + 'unedited_diffranges_cellcounts.tsv', 'w')
output.write('barcode\t'+'0-0.1'+'\t'+'0.1-0.2'+'\t'+'0.2-0.3'+'\t'+'0.3-0.4'+'\t'+'0.4-0.5'+'\t'+
             '0.5-0.6'+'\t'+'0.6-0.7'+'\t'+'0.7-0.8'+'\t'+'0.8-0.9'+'\t'+'0.9-1'+'\n')
for barcode in barcodes_dict.keys():  
        output.write(barcode+
                 '\t'+barcodes_dict[barcode]['0-0.1'][-1]+
                 '\t'+barcodes_dict[barcode]['0.1-0.2'][-1]+
                 '\t'+barcodes_dict[barcode]['0.2-0.3'][-1]+
                 '\t'+barcodes_dict[barcode]['0.3-0.4'][-1]+
                 '\t'+barcodes_dict[barcode]['0.4-0.5'][-1]+
                 '\t'+barcodes_dict[barcode]['0.5-0.6'][-1]+
                 '\t'+barcodes_dict[barcode]['0.6-0.7'][-1]+
                 '\t'+barcodes_dict[barcode]['0.7-0.8'][-1]+
                 '\t'+barcodes_dict[barcode]['0.8-0.9'][-1]+
                 '\t'+barcodes_dict[barcode]['0.9-1'][-1]+'\n')
output.close()
```

```{r make clonality dataframe}
clonality <- read.table("data/tables/unedited_diffranges_cellcounts.tsv", sep = "\t", header = T)
clonality$X0.0.1 <- as.vector(sapply(clonality$X0.0.1,normalize,length(subset(unedited@meta.data, barcode.count != 0 & diff.range == '0-0.1')$barcode)))
clonality$X0.1.0.2 <- as.vector(sapply(clonality$X0.1.0.2,normalize,length(subset(unedited@meta.data, barcode.count != 0 & diff.range == '0.1-0.2')$barcode)))
clonality$X0.2.0.3 <- as.vector(sapply(clonality$X0.2.0.3,normalize,length(subset(unedited@meta.data, barcode.count != 0 & diff.range == '0.2-0.3')$barcode)))
clonality$X0.3.0.4 <- as.vector(sapply(clonality$X0.3.0.4,normalize,length(subset(unedited@meta.data, barcode.count != 0 & diff.range == '0.3-0.4')$barcode)))
clonality$X0.4.0.5 <- as.vector(sapply(clonality$X0.4.0.5,normalize,length(subset(unedited@meta.data, barcode.count != 0 & diff.range == '0.4-0.5')$barcode)))
clonality$X0.5.0.6 <- as.vector(sapply(clonality$X0.5.0.6,normalize,length(subset(unedited@meta.data, barcode.count != 0 & diff.range == '0.5-0.6')$barcode)))
clonality$X0.6.0.7 <- as.vector(sapply(clonality$X0.6.0.7,normalize,length(subset(unedited@meta.data, barcode.count != 0 & diff.range == '0.6-0.7')$barcode)))
clonality$X0.7.0.8 <- as.vector(sapply(clonality$X0.7.0.8,normalize,length(subset(unedited@meta.data, barcode.count != 0 & diff.range == '0.7-0.8')$barcode)))
clonality$X0.8.0.9 <- as.vector(sapply(clonality$X0.8.0.9,normalize,length(subset(unedited@meta.data, barcode.count != 0 & diff.range == '0.8-0.9')$barcode)))
clonality$X0.9.1<- as.vector(sapply(clonality$X0.9.1,normalize,length(subset(unedited@meta.data, barcode.count != 0 & diff.range == '0.9-1.0')$barcode)))
names(clonality)[names(clonality) == "X0.0.1"] <- "0 - 0.1"
names(clonality)[names(clonality) == "X0.1.0.2"] <- "0.1 - 0.2"
names(clonality)[names(clonality) == "X0.2.0.3"] <- "0.2 - 0.3"
names(clonality)[names(clonality) == "X0.3.0.4"] <- "0.3 - 0.4"
names(clonality)[names(clonality) == "X0.4.0.5"] <- "0.4- 0.5"
names(clonality)[names(clonality) == "X0.5.0.6"] <- "0.5 - 0.6"
names(clonality)[names(clonality) == "X0.6.0.7"] <- "0.6 - 0.7"
names(clonality)[names(clonality) == "X0.7.0.8"] <- "0.7 - 0.8"
names(clonality)[names(clonality) == "X0.8.0.9"] <- "0.8 - 0.9"
names(clonality)[names(clonality) == "X0.9.1"] <- "0.9 - 1"
```

```{r plot line graph}
clonality_df <- melt(clonality, id.vars="barcode", value.name="value", variable.name= c("state"))
ggplot(data=clonality_df, aes(x=state, y=value, group = barcode, colour = barcode)) +
  #geom_point() + 
  geom_line(size = 1) +
  theme_bw() +
  theme(text = element_text(family = "Arial")) +
  theme(axis.text.y = element_text(size = 10)) +
  theme(axis.text.x = element_text(size = 10, angle = 25, hjust = 1)) +
  theme(axis.title = element_text(size = 12)) +
  theme(plot.title = element_text(size = 20)) +
  ggtitle("lineage barcode abundance across differentiation") + 
  theme(plot.title = element_text(hjust = 0.5, size = 12)) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("differentiation ranges") + 
  ylab("normalized percent") + ylim(0,100) + 
  scale_color_manual(values=rainbow(1)) + theme(legend.position = "none")
```

